Copulas and their potential for ecology

### 2019-10-07 Paper.Rmd
fig fixed: see response, please

<!--results for third mechanism-->
\begin{figure}[!h]
\textbf{ \hspace{3.5 cm} (A) \hspace{2 cm} (F) \hspace{2 cm} (K)} \\
\vspace{-1 cm}
\begin{center}
\includegraphics[width=9 cm]{./Results/BMR_results/evo_cop/ExtremeL_hist_taildep.pdf}\\
\end{center}
\vspace{-.33 cm}
\textbf{ \hspace{3.5 cm} (B) \hspace{2 cm} (G) \hspace{2 cm} (L)} \\
\vspace{-1 cm}
\begin{center}
\includegraphics[width=9 cm]{./Results/BMR_results/evo_cop/SomewhatL_hist_taildep.pdf}\\
\end{center}
\vspace{-.33 cm}
\textbf{ \hspace{3.5 cm} (C) \hspace{2 cm} (H) \hspace{2 cm} (M)} \\
\vspace{-1 cm}
\begin{center}
\includegraphics[width=9 cm]{./Results/BMR_results/evo_cop/BivNormal_hist_taildep.pdf}\\
\end{center}
\vspace{-.33 cm}
\textbf{ \hspace{3.5 cm} (D) \hspace{2 cm} (I) \hspace{2 cm} (N)} \\
\vspace{-1 cm}
\begin{center}
\includegraphics[width=9 cm]{./Results/BMR_results/evo_cop/SomewhatR_hist_taildep.pdf}\\
\end{center}
\vspace{-.33 cm}
\textbf{ \hspace{3.5 cm} (E) \hspace{2 cm} (J) \hspace{2 cm} (O)} \\
\vspace{-1 cm}
\begin{center}
\includegraphics[width=9 cm]{./Results/BMR_results/evo_cop/ExtremeR_hist_taildep.pdf}\\
\caption[Non-parametric statistics on a mechanism of tail dependence in character distributions]{Three 
measures of asymmetry of tail associations ($\cor_{0,0.2}-\cor_{0.8,1}$, A-E; 
$\Ps_{0,0.2}-\Ps_{0.8,1}$, F-J; $\Dsq_{0.8,1}-\Dsq_{0,0.2}$ K-O) between two characters, across extant species,  were 
computed for each of 100 simulations of mammalian character evolution for each of five types of tail dependence between 
evolutionary changes in the characters (extreme left-tail dependence (A, F, K), moderate left-tail dependence (B, G, L), 
symmetry of tail dependence (C, H, M), moderate right-tail dependence (D, I, N), and extreme right-tail dependence 
(E, J, O)). See the methods section for Q2 for details. The number above each panel indicates how many statistics out of $100$ were less than or equal to $0$. Values substantially lower (respectively, higher) than $50$ indicate greater left (respectively, right) tail-dependence between characters.\label{fig_cop_evo}}
\end{center}
\end{figure}
<!--***DAN: Shyamolina, can you please make this figure and its caption fit on one page,
somehow?-->

@@

\begin{figure}[!h]
\textbf{ \hspace{6.6 cm} (A) \hspace{4.6 cm} (B)} \\
\vspace{-0.4 cm}
\begin{center}
\includegraphics[width=5 cm]{./Results/skewness_results/skewness_aphid_count/skewness_aphid_count_sp_10_spearman.pdf}
\includegraphics[width=5 cm]{./Results/skewness_results/skewness_aphid_ff/skewness_aphid_ff_sp_11_spearman.pdf}\\
\end{center}
\vspace{0.2 cm}
\textbf{ \hspace{6.6 cm} (C) \hspace{4.6 cm} (D)} \\
\vspace{-0.4 cm}
\begin{center}
\includegraphics[width=5 cm]{./Results/skewness_results/skewness_plankton_north_sea/skewness_plankton_north_sea_sp_16_spearman.pdf}
\includegraphics[width=5 cm]{./Results/skewness_results/skewness_methane/skewness_methane_spearman.pdf}
\caption{Skewness of spatially averaged green spruce aphid abundance $(A)$, leaf-curling plum 
aphid first flight dates $(B)$, \emph{Ceratium furca} abundance $(C)$, and methane-flux
$(D)$ time series compared to a multivariate 
normal-copula null hypothesis. Black dots are empirical 
skewnesses; see text 
and Appendix \ref{SM-Surrog_multi} for details of the null hypothesis. Results show a 
tendency for skewness of the spatial average to be affected as hypothesized 
by asymmetric tail associations. \label{fig_skewness_spearman}}
\end{center}
\end{figure}

@@

\begin{figure}[!h]
\textbf{ \hspace{3.6 cm} (A) \hspace{4.2 cm} (B)} \\
\vspace{-0.8 cm}
\begin{center}
\includegraphics[width=10 cm]{./Results/ext_risk_copula/lcohen_ext_risk_local_disp.pdf}
\end{center}
\textbf{ \hspace{3.6 cm} (C) \hspace{4.2 cm} (D)} \\
\vspace{-0.8 cm}
\begin{center}
\includegraphics[width=10 cm]{./Results/ext_risk_copula/lcohen_ext_risk_global_disp.pdf}
\end{center}
\caption{Extinction risk for the metapopulation extension of the Lewontin-Cohen model, after 
$25$ time steps, was higher for environmental noise with stronger left-tail association
across space. Dispersal was local (A, B) or global (C, D),
with dispersal rate $d$ (Appendix \ref{SM-ext_risk_model}). Simulations used $N$ patches
for $N=5$ (A, C) and $N=25$ (B, D).\label{fig_extrisk_lcohen_local}}
\end{figure}

<!--adding plots showing influences of non-normal copula on spatial Taylors law-->
\begin{figure}[!h]
\textbf{ \hspace{1.4cm} (A) \hspace{2.92 cm} (B) \hspace{2.92 cm} (C) \hspace{2.92 cm} (D)} \\
\vspace{-0.5 cm} 
\begin{center}
\includegraphics[width=16 cm]{./Results/taylorslaw_results/SpatialTL_CNFSC_spearman.png}\\
\end{center}
\textbf{ \hspace{0.6 cm} (E) \hspace{1.4 cm} (F) \hspace{1.4 cm} (G) \hspace{1.4 cm} (H) \hspace{1.4 cm} (I) \hspace{1.4 cm} (J) \hspace{1.4 cm} (K)}\\
\vspace{-0.5 cm}
\begin{center}
\includegraphics[width=16 cm]{./Results/taylorslaw_results/SpatialTL_7stats_spearman.pdf}
\caption[Spatial Taylor's law results]{Spatial Taylor's law results. Log(spatial variance) vs. 
log(spatial mean) relationships for $1000$ simulations over $50$ time steps for $25$ locations 
using Clayton (A), normal (B), Frank (C), and survival Clayton (D) copula structures between
locations. Results from all $1000$ 
simulations were plotted on the same axes for each panel, but two example simulations are 
shown in red and green to help assess to what extent variation on the plots was between 
simulations or within simulations. For each simulation, we tested the linearity 
(E) and homoskedasticity (F) of the log(variance) vs. log(mean) plot for that simulation, 
and quantified the root mean squared error of points from the linear regression line (G).
We also calculated the linear regression intercept (H) and slope (I), the quadratic
term of a quadratic regression through the points on the log(variance) vs. log(mean) plot
(J), and the curvature of that quadratic regression (K). Distributions of values 
across all 1000 simulations are displayed. Spearman-preserving surrogates were used, 
though results using Kendall-preserving surrogates were similar. See the methods
section for Q3 and Appendices \ref{SM-Consequence_TL_simulation} and 
\ref{SM-sect:copsurrognd} for details.\label{fig_spTL}}
\end{center}
\end{figure}
<!--***DAN: Shyamolina, the spacing of the panel labels (A, B, C, etc.) is not quite 
matched up with the spacing of the panels themselves, can you fix it please?-->

@@

<div id="refs"></div>

>>>>

<!--results for third mechanism-->
\begin{figure}[!h]
\begin{center}
\includegraphics[width=15 cm]{./Results/BMR_results/evo_cop/singlepage_hist_taildep.pdf}\\
\caption[Non-parametric statistics on a mechanism of tail dependence in character distributions]{Three 
measures of asymmetry of tail associations ($\cor_{0,0.2}-\cor_{0.8,1}$, A-E; 
$\Ps_{0,0.2}-\Ps_{0.8,1}$, F-J; $\Dsq_{0.8,1}-\Dsq_{0,0.2}$ K-O) between two characters, across extant species,  were 
computed for each of 100 simulations of mammalian character evolution for each of five types of tail dependence between 
evolutionary changes in the characters (extreme left-tail dependence (A, F, K), moderate left-tail dependence (B, G, L), 
symmetry of tail dependence (C, H, M), moderate right-tail dependence (D, I, N), and extreme right-tail dependence 
(E, J, O)). See the methods section for Q2 for details. The number above each panel indicates how many statistics out of $100$ were less than or equal to $0$. Values substantially lower (respectively, higher) than $50$ indicate greater left (respectively, right) tail-dependence between characters.\label{fig_cop_evo}}
\end{center}
\end{figure}

@@

\begin{figure}[!h]
\textbf{ \hspace{5.6 cm} (A) \hspace{4 cm} (B)} \\
\vspace{-0.6 cm}
\begin{center}
\includegraphics[width=5 cm]{./Results/skewness_results/skewness_aphid_count/skewness_aphid_count_sp_10_spearman.pdf}
\includegraphics[width=5 cm]{./Results/skewness_results/skewness_aphid_ff/skewness_aphid_ff_sp_11_spearman.pdf}\\
\end{center}
\vspace{0.2 cm}
\textbf{ \hspace{5.6 cm} (C) \hspace{4 cm} (D)} \\
\vspace{-0.6 cm}
\begin{center}
\includegraphics[width=5 cm]{./Results/skewness_results/skewness_plankton_north_sea/skewness_plankton_north_sea_sp_16_spearman.pdf}
\includegraphics[width=5 cm]{./Results/skewness_results/skewness_methane/skewness_methane_spearman.pdf}
\caption{Skewness of spatially averaged green spruce aphid abundance (A), leaf-curling plum 
aphid first flight dates (B), \emph{Ceratium furca} abundance (C), and methane-flux
(D) time series compared to a multivariate 
normal-copula null hypothesis. Black dots are empirical 
skewnesses; see text 
and Appendix \ref{SM-Surrog_multi} for details of the null hypothesis. Results show a 
tendency for skewness of the spatial average to be affected as hypothesized 
by asymmetric tail associations. \label{fig_skewness_spearman}}
\end{center}
\end{figure}

@@

\begin{figure}[!h]
\textbf{ \hspace{5.6 cm} (A) \hspace{4 cm} (B)} \\
\vspace{-1 cm}
\begin{center}
\includegraphics[width=10 cm]{./Results/ext_risk_copula/lcohen_ext_risk_local_disp.pdf}
\end{center}
\textbf{ \hspace{5.6 cm} (C) \hspace{4 cm} (D)} \\
\vspace{-1 cm}
\begin{center}
\includegraphics[width=10 cm]{./Results/ext_risk_copula/lcohen_ext_risk_global_disp.pdf}
\end{center}
\caption{Extinction risk for the metapopulation extension of the Lewontin-Cohen model, after 
$25$ time steps, was higher for environmental noise with stronger left-tail association
across space. Dispersal was local (A, B) or global (C, D),
with dispersal rate $d$ (Appendix \ref{SM-ext_risk_model}). Simulations used $N$ patches
for $N=5$ (A, C) and $N=25$ (B, D).\label{fig_extrisk_lcohen_local}}
\end{figure}

<!--adding plots showing influences of non-normal copula on spatial Taylors law-->
\begin{figure}[!h]
\textbf{ \hspace{2cm} (A) \hspace{2.92 cm} (B) \hspace{2.92 cm} (C) \hspace{2.92 cm} (D)} \\
\vspace{-0.8 cm} 
\begin{center}
\includegraphics[width=16 cm]{./Results/taylorslaw_results/SpatialTL_CNFSC_spearman.png}\\
\end{center}
\textbf{ \hspace{1 cm} (E) \hspace{1.4 cm} (F) \hspace{1.2 cm} (G) \hspace{1.2 cm} (H) \hspace{1.4 cm} (I) \hspace{1.4 cm} (J) \hspace{1.2 cm} (K)}\\
\vspace{-0.8 cm}
\begin{center}
\includegraphics[width=16 cm]{./Results/taylorslaw_results/SpatialTL_7stats_spearman.pdf}
\caption[Spatial Taylor's law results]{Spatial Taylor's law results. Log(spatial variance) vs. 
log(spatial mean) relationships for $1000$ simulations over $50$ time steps for $25$ locations 
using Clayton (A), normal (B), Frank (C), and survival Clayton (D) copula structures between
locations. Results from all $1000$ 
simulations were plotted on the same axes for each panel, but two example simulations are 
shown in red and green to help assess to what extent variation on the plots was between 
simulations or within simulations. For each simulation, we tested the linearity 
(E) and homoskedasticity (F) of the log(variance) vs. log(mean) plot for that simulation, 
and quantified the root mean squared error of points from the linear regression line (G).
We also calculated the linear regression intercept (H) and slope (I), the quadratic
term of a quadratic regression through the points on the log(variance) vs. log(mean) plot
(J), and the curvature of that quadratic regression (K). Distributions of values 
across all 1000 simulations are displayed. Spearman-preserving surrogates were used, 
though results using Kendall-preserving surrogates were similar. See the methods
section for Q3 and Appendices \ref{SM-Consequence_TL_simulation} and 
\ref{SM-sect:copsurrognd} for details.\label{fig_spTL}}
\end{center}
\end{figure}
<!--***DAN: Shyamolina, the spacing of the panel labels (A, B, C, etc.) is not quite 
matched up with the spacing of the panels themselves, can you fix it please?
***Shya: check now, if is it ok?-->

@@

<!--<div id="refs"></div>-->
